---
title: "Statistical Inference Course Project"
author: "nthehai01"
date: "2/23/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Overview

The project consists of two parts:

1. A simulation exercise
2. Basic inferential data analysis


## Part 1: Simulation Exercise

#### Simulations

In this project, I will investigate the exponential distribution of averages of 40 exponentials with $\lambda$ = 0.2 and compare it with the Central Limit Theorem.
```{r}
## initialize
n = 40
lambda = 0.2
```

Creating a thousand exponential distribution simulations:
```{r}
n.sim = 1000

set.seed(42)
sim <- replicate(n.sim, rexp(n, lambda))
```

Calculating mean of the each simulation:
```{r}
mean.sim <- apply(sim, 2, mean)
```


#### Sample Mean versus Theoretical Mean

**Note:** the theoretical mean and standard deviation of exponential distribution are the same with $$\frac{1}{\lambda}$$.

The sample mean:
```{r}
mean.sam <- mean(mean.sim)
mean.sam
```

The theoretical mean:
```{r}
mean.theo <- 1 / lambda
mean.theo
```

As can be seen from the results, the sample mean and the theoretical mean are pretty much close.


#### Sample Variance versus Theoretical Variance

The sample variance:
```{r}
var.sam <- var(mean.sim)
var.sam
```

The theoretical variance:
```{r}
var.theo <- (1/lambda)^2 / n
var.theo
```

As can be seen from the results, the sample variance and the theoretical variance are pretty much close.


#### Distribution

In this session, I will plot the simulations and two distribution curves, one for the simulations and one for the theoretical distribution.

```{r}
mean.df <- data.frame(mean = mean.sim)
```

Creating a histogram for the simulations:
```{r}
library(ggplot2)
hist.sim <- ggplot(mean.df, aes(x = mean)) +
  geom_histogram(aes(y =..density..),
                 binwidth = lambda,
                 fill="lightcoral", 
                 color="black") +
  ggtitle("Exponential distribution")
```

Adding a normal curve for the simulations:
```{r}
hist.sim <- hist.sim + stat_function(fun = dnorm, 
                                     args = list(mean = mean.sam, 
                                                 sd = sqrt(var.sam)),
                                     aes(colour = "Simulations"),
                                     size = 1.5)
```

Adding a normal curve for the theoritial distribution:
```{r}
hist.sim <- hist.sim + stat_function(fun = dnorm, 
                                     args = list(mean = mean.theo, 
                                                 sd = sqrt(var.theo)),
                                     aes(colour = "Theory"),
                                     alpha = 0.6,
                                     size = 1.5)
```

Adding color to two distribution curves:
```{r}
hist.sim <- hist.sim + scale_colour_manual("Type", values=c("yellow", "black"))
```

Plotting:
```{r}
hist.sim
```

As can be seen from the graph, the two distributions are identical.

## Part 2

In the second portion of the project, I am going to analyze the ToothGrowth data in the R datasets package.

#### Loading the ToothGrowth data 
```{r}
data(ToothGrowth)
```


#### Performing some basic exploratory data analyses

The response is the *length* of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three *dose* levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery *methods*, orange juice (coded as OJ) or ascorbic acid (a form of vitamin C and coded as VC). 

```{r}
str(ToothGrowth)
```

It shows that our data set containing 60 observations and 3 variables with ```len``` and ```dose``` are numerical, ```supp``` is a 2-level factor.

Check for missing values:
```{r}
apply(is.na(ToothGrowth), 2, sum)
```

Let's look at how many samples we have for each type of doe and method pairs:
```{r}
table(ToothGrowth$supp, ToothGrowth$dose)
```

We see that there are 10 samples for each pair of dose and method, 20 samples for each dose and 30 samples for each method.

Let's look at a box plot of tooth length broken down by method and dose:
```{r}
library(ggplot2)
ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = supp)) +
  geom_boxplot() +
  ggtitle("Tooth length by method and dose") +
  xlab("Dose (mg/day)") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title='Method'))
```

From the plot, we see:

1. For 2.0 mg/day dose, there is no difference in tooth length between OJ and VC.
2. For 0.5 mg/day and 1.0 mg/day dose, the lengths of odontoblasts with OJ method are higher than with VC method.
3. For OJ and VC, a dose of 1.0 mg/day is always higher than a dose of 0.5 mg/day when it comes to tooth length.


#### Providing a basic summary of the data

```{r}
summary(ToothGrowth)
```


#### Using confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose

**First hypothesis:** *$H_0$:* Both types of supplement are involving in tooth length.
```{r}
t.test(len ~ supp, data = ToothGrowth, alternative = "two.sided")
```

Since the p-value > $\alpha$ = 0.05, we can not reject the null hypothesis. Both supplements are delivering same tooth growth.

**Second hypothesis:** *$H_0$:* For 0.5 mg/day and 1.0 mg/day dose, the lengths of odontoblasts with OJ method are less than or equal to VC method.
```{r}
## filter the record with 0.5 or 1.0 dose
filtered.dose <- ToothGrowth[ToothGrowth$dose != 2.0, ]

t.test(len ~ supp, data = filtered.dose, alternative = "greater")
```

Since the p-value < $\alpha$ = 0.05, we accept the null hypothesis that for 0.5 mg/day and 1.0 mg/day dose, the lengths of odontoblasts with OJ method are less than or equal to VC method.

**Third hypothesis:** *$H_0$:* For OJ and VC, a dose of 0.5 mg/day is always higher than or equal to a dose of 1.0 mg/day when it comes to tooth length.
```{r}
t.test(len ~ dose, data = filtered.dose, alternative = "less")
```

Since the p-value < $\alpha$ = 0.05, we reject the null hypothesis. For OJ and VC supplements, a dose of 1.0 mg/day is always higher than a 0.5 mg/day dose in term of tooth length.


#### Conclusion

According to the t-tests applied in the previous section and according to the summarised data, it is clear that:

- The OJ and VC supplements are delivering same tooth growth but for 0.5 mg/day and 1.0 mg/day dose, the lengths of odontoblasts with OJ method are greater than VC method. 
- For OJ and VC supplements, a dose of 1.0 mg/day is always higher than a 0.5 mg/day dose in term of tooth length.
