---
title: "Statistical Inference Course Project"
author: "nthehai01"
date: "2/23/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Overview

The project consists of two parts:

1. A simulation exercise
2. Basic inferential data analysis


## Part 1: Simulation Exercise

#### Simulations

In this project, I will investigate the exponential distribution of averages of 40 exponentials with $\lambda$ = 0.2 and compare it with the Central Limit Theorem.
```{r}
## initialize
n = 40
lambda = 0.2
```

Creating a thousand exponential distribution simulations:
```{r}
n.sim = 1000

set.seed(42)
sim <- replicate(n.sim, rexp(n, lambda))
```

Calculating mean of the each simulation:
```{r}
mean.sim <- apply(sim, 2, mean)
```


#### Sample Mean versus Theoretical Mean

**Note:** the theoretical mean and standard deviation of exponential distribution are the same with $$\frac{1}{\lambda}$$.

The sample mean:
```{r}
mean.sam <- mean(mean.sim)
mean.sam
```

The theoretical mean:
```{r}
mean.theo <- 1 / lambda
mean.theo
```

As can be seen from the results, the sample mean and the theoretical mean are pretty much close.


#### Sample Variance versus Theoretical Variance

The sample variance:
```{r}
var.sam <- var(mean.sim)
var.sam
```

The theoretical variance:
```{r}
var.theo <- (1/lambda)^2 / n
var.theo
```

As can be seen from the results, the sample variance and the theoretical variance are pretty much close.


#### Distribution

In this session, I will plot the simulations and two distribution curves, one for the simulations and one for the theoretical distribution.

```{r}
mean.df <- data.frame(mean = mean.sim)
```

Creating a histogram for the simulations:
```{r}
library(ggplot2)
hist.sim <- ggplot(mean.df, aes(x = mean)) +
  geom_histogram(aes(y =..density..),
                 binwidth = lambda,
                 fill="lightcoral", 
                 color="black") +
  ggtitle("Exponential distribution")
```

Adding a normal curve for the simulations:
```{r}
hist.sim <- hist.sim + stat_function(fun = dnorm, 
                                     args = list(mean = mean.sam, 
                                                 sd = sqrt(var.sam)),
                                     aes(colour = "Simulations"),
                                     size = 1.5)
```

Adding a normal curve for the theoritial distribution:
```{r}
hist.sim <- hist.sim + stat_function(fun = dnorm, 
                                     args = list(mean = mean.theo, 
                                                 sd = sqrt(var.theo)),
                                     aes(colour = "Theory"),
                                     alpha = 0.6,
                                     size = 1.5)
```

Adding color to two distribution curves:
```{r}
hist.sim <- hist.sim + scale_colour_manual("Type", values=c("yellow", "black"))
```

Plotting:
```{r}
hist.sim
```

As can be seen from the graph, the two distributions are identical.

## Part 2

